---
title: "TBD"
author: "Smirnova, E. et al"
date: "`r Sys.Date()`"
output:
  html_document
---


```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
rm(list=ls())
# Set up the environment
library(knitr)
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700, 
library(pander)
panderOptions('table.split.table', Inf)
set.seed(1)
library(dplyr)
options(stringsAsFactors = FALSE)
set.seed(41235)
```


```{r libraries, echo=FALSE}
library(ggplot2)
library(readxl)
library(kableExtra)
library(cowplot)
library(plyr)
library(dplyr)
#library(magrittr)
library(reshape2)
library(tableone)
library(caret)
library(plotly)
require(tidyverse)
require(tidymodels)
require(data.table)
require(tidyposterior)
require(tsibble)  #tsibble for time series based on tidy principles
require(fable)  #for forecasting based on tidy principles
require(ggfortify)  #for plotting timeseries
require(forecast)  #for forecast function
require(tseries)
require(chron)
require(lubridate)
require(directlabels)
require(zoo)
require(lmtest)
require(TTR)  #for smoothing the time series
require(MTS)
require(vars)
require(fUnitRoots)
require(lattice)
require(grid)


#https://bookdown.org/singh_pratap_tejendra/intro_time_series_r/multivariate-ts-analysis.html
```

```{r settings, echo=FALSE}
mainDir <- here::here("Ethereum")

# list of token files for betti numbers
bettiDir <- file.path(mainDir, "betti")
B0 <- read_rds(file.path(bettiDir, "allTokens_B0.rds"))
B1 <- read_rds(file.path(bettiDir, "allTokens_B1.rds"))

#rolling depth files
depthDir <- file.path(mainDir, "depth")
B0_depth <- read_rds(file.path(depthDir, "rd_allTokens_B0.rds"))
B1_depth <- read_rds(file.path(depthDir, "rd_allTokens_B1.rds"))

#merged data
mergeDir <- file.path(mainDir, "merge")
threshold <- 0.15
data <- read_rds(file.path(mergeDir, paste0("df_allTokens_abs", 100*threshold, ".rds")))

```

# Classification tables for each token

rows are predicted, columns are actual

anomaly threshold = 0.25

prediction horizon = 3

prediction length = 100 

```{r}
df.m <- melt(data[c("day", "B0", "B1", "priceReturn")], id.vars = "day")
p <- ggplot(df.m, aes(x=day, y = value, color = variable))+
  geom_line()+
  facet_grid(rows = "variable", scales= "free")+
  theme_bw()+
  theme(legend.position = "none")

#ggplotly(p)
```


```{r}
#################################
#price prediction using multivariate TS
##################################
vars_predict <- c("B0", "B1", "priceReturn")
inx <- apply(is.na(data[, which(names(data) %in% vars_predict)]), 1, any)
df.no.na <- data[!inx, ]

#prediction horizon
pred_h = 3
pred_length = 99#actually 100 but loop for i starts from 1
total_length= dim(df.no.na)[1]

#create data frame to store qulity results for each i
quality <- data.frame(actual_anom=logical(),
                      pred_anom=logical(),
                     stringsAsFactors=FALSE)
#to hold model fit results for each model
MV_TSlist <- list()
for (j in 1: (total_length - pred_h -pred_length)){
  df_pred <- df.no.na[1:(pred_length +j),vars_predict]
  newdata <- df.no.na[(pred_length +j+1):(pred_length +j+pred_h),]
  # Converting a data.frame into mts
  mymts = ts(df_pred,
           frequency=1,
           start = 1)

  #######################################################
  #time series prediction
  #######################################################

  # Creating a VAR model with vars
  var.a <- vars::VAR(mymts,
                   lag.max = 10, #highest lag order for lag length selection according to the choosen ic
                   ic = "AIC", #information criterion
                   type = "none") #type of deterministic regressors to include
  #predict next h days
  pred.i <- predict(var.a, newdata, n.ahead = pred_h)
  #compare with actual
  pred.price <- pred.i$fcst$priceReturn[,which(colnames(pred.i$fcst$priceReturn) == "fcst")]
  actual_anom <- max(abs(newdata$priceReturn)) >=threshold
  pred_anom <- max(abs(pred.price)) >=threshold
  quality[j,] <- c(actual_anom, pred_anom)
  
  #############################
  #model fit for each forecast
  ############################
  p_vals <- sapply(summary(var.a)$varresult, function(x) {
    tmp <- x[["fstatistic"]]
     pf(tmp[1], tmp[2], tmp[3], lower.tail = FALSE)  })

  r_sq <- lapply(summary(var.a)$varresult, "[", "adj.r.squared") 
  r_sq <- unlist(sapply(r_sq, "["))
  res <- data.frame(p_val = p_vals,
                  R_sq_adj = r_sq)
  rownames(res) <- c("B0", "B1", "priceReturn")
  #######################################################
  #gragner causality
  #######################################################

  #selecting the variables
  # Granger test for causality
   #for causality function to give reliable results we need all the variables of the multivariate time series to be stationary. 
  GC <- causality(var.a, #VAR model
          cause = c("B0", "B1"))

   Causal <-c(GC$Granger$p.value, GC$Instant$p.value) #h0: do not cause: large p-value -- no relationship
   names(Causal) <- c("Granger", "Instant")

   res_ls <- list(
              TS = res,
              Causality = Causal)

   MV_TSlist[[j]] <- res_ls
  
}#end for j


```

```{r}
classif <- table(pred_anom = quality$pred_anom, actual_anom = quality$actual_anom)
classif
```


Multivariate time series results: model fit

```{r, echo=FALSE}
#p-values for each token
df <- sapply(MV_TSlist,  function(x) x$TS$p_val)
rownames(df) <- c("B0", "B1", "priceReturn")

kable(df, digits = 3, caption = "Model p-values")%>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
```

Multivariate time series results: adjusted r-squared

```{r, echo=FALSE}
#p-values for each token
df <- sapply(MV_TSlist,  function(x) x$TS$R_sq_adj)
rownames(df) <- c("B0", "B1", "priceReturn")

kable(df, digits = 3, caption = "Model R-squared")%>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
```


Multivariate time series results: gragner causality p-value

H0: B0 and B1 do not cause price return: large p-value -- no relationship

```{r, echo=FALSE}
#p-values for each token
df <- sapply(MV_TSlist,  function(x) x$Causality)

kable(df, digits = 3, caption = "Model Gragner Causality p-values")%>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
```

# Regression-based approach

Predict flag 1 to 7 

```{r}

library(mgcv)

vars_predict <- c("B0", "B1", "priceReturn")
inx <- apply(is.na(data[, which(names(data) %in% vars_predict)]), 1, any)
df.no.na <- data[!inx, ]

data_train <- df.no.na[1:300,]
data_test <- df.no.na[301:dim(df.no.na)[1],]

#logistic regression predictions
fit.gam <- gam(flag7 ~ vertexNum + edgeNum + clusterCoef + 
      s(B0, bs = "ps", k = 7)+
      s(B1, bs = "ps", k = 7),
    family = binomial(link= "logit"),
    data = data_train)

summary(fit.gam)

#confusion matrix

```

Non-linear effect in betti 0 

```{r}
layout(matrix(1:2, nrow = 1))
plot(fit.gam, shade = TRUE)
```

Predictions

```{r}

predicted <- predict(fit.gam, type = "response")
df.pred <- data.frame(actual = ifelse(data_train$flag7 == TRUE, 1, 0),
                      predicted = predicted)

#in your case, this will be 
df.pred$actual <- factor(df.pred$actual)
hist_predicted = ggplot(df.pred, 
                        aes(predicted, group=actual, fill = actual)) + 
  geom_histogram(aes(y=..density..),
                 position="identity", bins = 20, alpha = 0.6)
hist_predicted
```


```{r}
auc <- roc(response = df.pred$actual, predictor = df.pred$predicted)
auc
p.roc <- ggroc(auc, color = "darkred") 

p.roc <- p.roc +
  annotate("text", x = 0.25, y = 0.75, label = paste("AUC =", round(auc$auc[1],3)),
           color = "darkred", size = 6)+
  theme_bw()+  theme(axis.title.x = element_text( size=12), 
                     axis.text.x  = element_text( size=12),
                     axis.text.y  = element_text( size=12),
                     axis.title.y=element_text(size=12))
p.roc
```


Confusion matrix: training

```{r}
table(actual = df.pred$actual,
      predicted = ifelse(df.pred$predicted <0.5, 0,1))
```


Predictions test

```{r}

predicted <- predict(fit.gam, newdata = data_test, type = "response")
df.pred <- data.frame(actual = ifelse(data_test$flag7 == TRUE, 1, 0),
                      predicted = predicted)

#in your case, this will be 
df.pred$actual <- factor(df.pred$actual)
hist_predicted = ggplot(df.pred, 
                        aes(predicted, group=actual, fill = actual)) + 
  geom_histogram(aes(y=..density..),
                 position="identity", bins = 20, alpha = 0.6)
hist_predicted
```


```{r}
auc <- roc(response = df.pred$actual, predictor = df.pred$predicted)
auc
p.roc <- ggroc(auc, color = "darkred") 

p.roc <- p.roc +
  annotate("text", x = 0.25, y = 0.75, label = paste("AUC =", round(auc$auc[1],3)),
           color = "darkred", size = 6)+
  theme_bw()+  theme(axis.title.x = element_text( size=12), 
                     axis.text.x  = element_text( size=12),
                     axis.text.y  = element_text( size=12),
                     axis.title.y=element_text(size=12))
p.roc
```


Confusion matrix: training

```{r}
table(actual = df.pred$actual,
      predicted = ifelse(df.pred$predicted <0.5, 0,1))
```

# Gam with interaction between betti 0 and betti 1


```{r}

#logistic regression predictions
fit.gam <- gam(flag7 ~ vertexNum + edgeNum + clusterCoef + 
      s(B0, B1, bs = "tp", k = 7),
    family = binomial(link= "logit"),
    data = data_train)

summary(fit.gam)
```

# Gam with autocorrelation

```{r}

gam_6_ar0 <- gamm(flag7 ~ t2(B0, B1,
                            k = 7,
                            bs = c("ps", "ps"),
                            full = TRUE),
                  data = data_train,
                  family = binomial(link= "logit"),
                  method = "REML")

gam_7_ar0 <- gamm(flag7 ~ vertexNum + edgeNum + clusterCoef + 
                  s(B0, bs = "ps", k = 7)+
                  s(B1, bs = "ps", k = 7),
                  data = data_train,
                  family = binomial(link= "logit"),
                  method = "REML")
 
gam_6_ar1 <- gamm(flag7 ~ vertexNum + edgeNum + clusterCoef + 
                  s(B0, bs = "ps", k = 7)+
                  s(B1, bs = "ps", k = 7),
                  correlation = corARMA(form = ~ 1|B0, p = 1),
             data = data_train,
                  family = binomial(link= "logit"),
                  method = "REML")
```

