---
title: "TBD"
author: "Smirnova, E. et al"
date: "`r Sys.Date()`"
output:
  html_document
---


```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
rm(list=ls())
# Set up the environment
library(knitr)
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700, 
library(pander)
panderOptions('table.split.table', Inf)
set.seed(1)
library(dplyr)
options(stringsAsFactors = FALSE)
set.seed(41235)
```


```{r libraries, echo=FALSE}
library(ggplot2)
library(readxl)
library(kableExtra)
library(cowplot)
library(plyr)
library(dplyr)
#library(magrittr)
library(reshape2)
library(tableone)
library(caret)
library(plotly)
require(tidyverse)
require(tidymodels)
require(data.table)
require(tidyposterior)
require(tsibble)  #tsibble for time series based on tidy principles
require(fable)  #for forecasting based on tidy principles
require(ggfortify)  #for plotting timeseries
require(forecast)  #for forecast function
require(tseries)
require(chron)
require(lubridate)
require(directlabels)
require(zoo)
require(lmtest)
require(TTR)  #for smoothing the time series
require(MTS)
require(vars)
require(fUnitRoots)
require(lattice)
require(grid)
library(mgcv)

#https://bookdown.org/singh_pratap_tejendra/intro_time_series_r/multivariate-ts-analysis.html
```

```{r settings, echo=FALSE}
mainDir <- here::here("Ethereum")

#merged data
mergeDir <- file.path(mainDir, "merge")
threshold <- 0.15
data <- read_rds(file.path(mergeDir, paste0("df_allTokens_abs", 100*threshold, ".rds")))

```


# Regression-based approach using non-linear effects in betti 0 and betti 1


```{r}
runGam <- function(df, nflags, inputfmls){
  
#create holder for results
accuracy_df = data.frame(matrix(vector(), nflags, length(inputfmls),
                dimnames=list(c(), names(inputfmls))),
                stringsAsFactors=F)

auc_df <- data.frame(matrix(vector(), nflags, length(inputfmls),
                dimnames=list(c(), names(inputfmls))),
                stringsAsFactors=F)

sensitivity_df <- data.frame(matrix(vector(), nflags, length(inputfmls),
                dimnames=list(c(), names(inputfmls))),
                stringsAsFactors=F)

specificity_df <- data.frame(matrix(vector(), nflags, length(inputfmls),
                dimnames=list(c(), names(inputfmls))),
                stringsAsFactors=F)

precision_df <- data.frame(matrix(vector(), nflags, length(inputfmls),
                dimnames=list(c(), names(inputfmls))),
                stringsAsFactors=F)

aic_lst <- list()
pvals_lst <- list()

  for(j in 1:length(inputfmls)){

    fmlname <- names(inputfmls)[j]
    fml <- inputfmls[j]
    cftb_lst<- list()#to store model accuracy results
    aic_vec <-c()
      for(i in 1:nflags){
        flagi <- paste0("flag", i)
        vars_predict <- c("B0", "B1")

        flags = df %>% dplyr::select(contains("flag")) %>% colnames()
        df = df %>% mutate_if(is.logical,factor,levels=c("FALSE", "TRUE"))
        df$clusterCoef = replace(df$clusterCoef,is.nan(df$clusterCoef),0)
        dateRange = range(df$day)
        spDate = dateRange[1] + diff(dateRange)/3*2
        dftrain = df[df$day<spDate,]
        dftest = df[!df$day<spDate,]

        inx <- apply(is.na(dftrain[, which(names(dftrain) %in% vars_predict)]), 1, any)
        dftrain <- dftrain[!inx, ]  
        fm <- as.formula(paste(flagi, fml))
        #logistic regression predictions
        fit.gam <- gam(fm,
        family = binomial(link= "logit"),
            data = dftrain)

        aic_vec <- c(aic_vec, AIC(fit.gam))
        #p-value for coefficients
          if(i ==1){
            p_val <- c(summary(fit.gam)$p.pv, summary(fit.gam)$s.pv)
            names(p_val) <- c(rownames(summary(fit.gam)$p.table), rownames(summary(fit.gam)$s.table))}
          else{p_val <- rbind(p_val, c(summary(fit.gam)$p.pv, summary(fit.gam)$s.pv))}

        predicted <- ifelse(predict(fit.gam,dftest, type = "response") <0.5, "FALSE", "TRUE")
        predicted <- factor(predicted, levels = c("TRUE", "FALSE"))
        cf = caret::confusionMatrix(
            predicted,
            dftest %>% pull(flagi),
            positive = "TRUE"
          )

        cfmatrix = rbind(as.matrix(cf,what="overall"),
                           as.matrix(cf,what="classes"))
        cftb = cfmatrix %>% as.data.frame()  %>% rownames_to_column() %>% as_tibble()
        names(cftb) = c("measure","value")
        cftb[nrow(cftb)+1,"measure"] <- "AUC"
        cftb[nrow(cftb),"value"] <- pROC::auc(unlist(dftest[,flagi]),
                          predict(fit.gam,dftest, type = "response"),
                          levels = c("FALSE", "TRUE"), quiet=T) %>% as.numeric()
        cftb$horizon = flagi
        cftb$modelType = fmlname
        #save as a list for each flag
        cftb_lst[[i]] <- cftb
      }

    #save aic and p-vals for each model
    aic_lst[[j]] <- aic_vec
    pvals_lst[[j]] <- p_val


    #extract accuracy for each flag
    accuracy_df[,j] <- unlist(lapply(cftb_lst, function(x) x[x$measure == "Accuracy", "value"]))
    auc_df[,j] <- unlist(lapply(cftb_lst, function(x) x[x$measure == "AUC", "value"]))
    sensitivity_df[,j] <- unlist(lapply(cftb_lst, function(x) x[x$measure == "Sensitivity", "value"]))
    specificity_df[,j] <- unlist(lapply(cftb_lst, function(x) x[x$measure == "Specificity", "value"]))
    precision_df[,j] <- unlist(lapply(cftb_lst, function(x) x[x$measure == "Precision", "value"]))

  }#end for j

accuracy_df$horizon <- paste0("flag", 1:nflags)
auc_df$horizon <- paste0("flag", 1:nflags)
sensitivity_df$horizon <- paste0("flag", 1:nflags)
specificity_df$horizon <- paste0("flag", 1:nflags)
precision_df$horizon <- paste0("flag", 1:nflags)

return(list(aic = aic_lst, pvals = pvals_lst,
            accuracy = accuracy_df,
            auc = auc_df,
            sensitivity = sensitivity_df,
            specificity = specificity_df,
            precision = precision_df))
}
```


## Models without autocorrelation

For each of the flags 1, ... 7, the following models are fitted

M1= ~ vertexNum + edgeNum + clusterCoef + openNorm

M2=~ vertexNum + edgeNum + clusterCoef + openNorm + s(B0, bs = 'ps', k = 7)

M3=~ vertexNum + edgeNum + clusterCoef + openNorm + s(B0, bs = 'ps', k = 7) + s(B0, bs = 'ps', k = 7)

```{r}
inputfmls = c(M1="~ vertexNum + edgeNum + clusterCoef + openNorm",
                M2="~ vertexNum + edgeNum + clusterCoef + openNorm + s(B0, bs = 'ps', k = 7)",
                M3="~ vertexNum + edgeNum + clusterCoef + openNorm + s(B0, bs = 'ps', k = 7) + s(B0, bs = 'ps', k = 7)")

res_nocorr <- runGam(df = data, nflags = 7, inputfmls = inputfmls)

```

AUC for each model

```{r}
kable(res_nocorr$auc, digits =3)%>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
```

Accuracy for each model

```{r}
kable(res_nocorr$accuracy, digits =3)%>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
```

Sensitivity for each model

```{r}
kable(res_nocorr$sensitivity, digits =3)%>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
```

Specificity for each model

```{r}
kable(res_nocorr$specificity, digits =3)%>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
```

Precision for each model

```{r}
kable(res_nocorr$precision, digits =3)%>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
```


Non-linear effect in betti 0 

```{r, eval = FALSE}
layout(matrix(1:2, nrow = 1))
plot(fit.gam, shade = TRUE)
```


# Gam with interaction between betti 0 and betti 1


```{r, eval = FALSE}

#logistic regression predictions
fit.gam <- gam(flag7 ~ vertexNum + edgeNum + clusterCoef + 
      s(B0, B1, bs = "tp", k = 7),
    family = binomial(link= "logit"),
    data = data_train)

summary(fit.gam)
```

# Gam with autocorrelation

```{r, eval = FALSE}

gam_6_ar0 <- gamm(flag7 ~ t2(B0, B1,
                            k = 7,
                            bs = c("ps", "ps"),
                            full = TRUE),
                  data = data_train,
                  family = binomial(link= "logit"),
                  method = "REML")

gam_7_ar0 <- gamm(flag7 ~ vertexNum + edgeNum + clusterCoef + 
                  s(B0, bs = "ps", k = 7)+
                  s(B1, bs = "ps", k = 7),
                  data = data_train,
                  family = binomial(link= "logit"),
                  method = "REML")
 
gam_6_ar1 <- gamm(flag7 ~ vertexNum + edgeNum + clusterCoef + 
                  s(B0, bs = "ps", k = 7)+
                  s(B1, bs = "ps", k = 7),
                  correlation = corARMA(form = ~ 1|B0, p = 1),
             data = data_train,
                  family = binomial(link= "logit"),
                  method = "REML")
```

